---
title: "Statistical Computing: Assignment 1"
subtitle: "Investigating Globular Clusters and Galaxies"
author: "Louwrens Labuschagne - LBSLOU003"
institute: "Department of Statistical Sciences, University of Cape Twon"
date:  "`r format(Sys.time(), '%d %B %Y')`"
output: pdf_document
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
 
```

## Introduction

### Background
The catalog we will be investigating (@catalog) is based on a literature survey to the end of 2012 and consists of 422 galaxies with published measurements of their globular cluster (GC) populations. The galaxy morphological classification of these consist of 248 E galaxies, 93 S0 galaxies, and 81 spirals or irregulars. [@1_harris_harris_alessi_2013] To gain some insight into these morphological classification types we take a look at how the types look in Figure \ref{fig:HubbleTuningFork}, @hubbleClassificationScheme.

``` {r eval=T, echo=F, plot-ref, fig.cap = "Galaxy Morphological Classification a.k.a. Hubble Tuning Fork\\label{fig:HubbleTuningFork}", echo = FALSE}
```
\begin{figure}[h]
	\label{fig:HubbleTuningFork}
	\center
	\includegraphics[width=0.7\textwidth]{assets/HubbleTuningFork.jpg}
	\caption{Galaxy Morphological Classification a.k.a. Hubble Tuning Fork}
\end{figure}


## Data Gathering And Cleaning


```{r echo=F, eval=T, code = readLines("dataCleaning.R")}
```

## Analysis
### central black hole mass

### dynamical bulge mass

### bulge velocity dispersion

### absolute visual magnitude and type of galaxy

## Conclusions
Look at NGC221 M32 as N(GC) = 0 - An exception is the Local Group elliptical M32, which we include for historical reasons though it has no clearly identified clusters.

## References